|

### 2024-02-02 rmds/manuscript.Rmd
manuscript

---
title: |
  Quantifying maternal reproductive output of chondrichthyan fishes
subtitle: |
author:
- name: Alastair V Harry
  affiliation: Fisheries & Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia
  email: alastair.harry@gmail.com
  number: 1
- name: Ivy Baremore
  affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
  number: 2
- name: Andrew Piercy
  affiliation: College of Southern Maryland, La Plata MD 20646, United States
  number: 3
abstract: |
  For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to calculate the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, relatively few studies report maternity functions. Instead, maturity functions have typically been used as a proxy, despite evidence of a delay between attainment of functional maturity and pregnancy in some species. This study examined the relative performance of alternative approaches for quantifying maternal reproductive output. Applying logistic models to a combination of simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used. At sample sizes < 250 individuals a fixed-value for $P_{Max}$ was more effective at reducing bias in simulated data for gummy sharks, *Mustelus antarcticus*. Maturity parameters could be estimated with greater precision however substituting them in place of maternity parameters overestimated lifetime reproductive output. The extent to which maturity functions can provide a good approximation for maternity functions may be species specific and requires futher research. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. We proposed that estimation of $P_{Max}$ is generally preferable to using a fixed value. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.

toc: false
output:
  bookdown::pdf_document2:
    pandoc_args:
      - --csl=extras/cjfas.csl
      - --citation-abbreviations=extras/abbreviations.json
      - --template=extras/mytemplate.tex
bibliography: "extras/maternal-output.bib"
keep-latex: true
linestretch: 2 
header-includes:
  - \fontsize{12}{20}
  - \usepackage{lineno}
  - \linenumbers
  - \linenumberfont
  - \renewcommand\linenumberfont{\normalfont\small\sffamily}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{\thepage}
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
editor_options: 
  chunk_output_type: inline
---

@@

@@

Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass [@quinn_quantitative_1999]. Calculating reproductive capacity requires data on the sexually mature proportion of the population contributing to reproduction at a given size or age. For most populations of teleost fishes a simple two-parameter logistic regression function (2PLF) is sufficient to model this process using dichotomous sexual maturity stage data (immature / mature) from a sample of individuals [@jennings_marine_2001; @king_fisheries_2007]. While the mathematical formulae and methods of statistically estimating parameters for a 2PLF vary, the estimated values are similar and can be obtained with the built in generalised linear model (GLM) programs in most modern statistical software.

In chondrichthyan fishes (sharks, rays, skates, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015]. Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017]. This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014]. Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stockrecruitment_2013]. This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

@@

# Methods

@@

A logical progression from estimating $P_{Max}$ is the consideration of alternative functional forms for this parameter such as time- or space-varying $P_{Max}$ or more complex length- or age-dependent forms. Despite the diverse range of reproductive modes found in chondrichthyan fishes, the current, limited knowledge of maternal investment has typically investigated variables such as litter size and pup condition [@hussey_maternal_2010]. Little is known about how reproductive frequency varies as a function of size or age, in response to environmental effects, or at changing population densities. Better understanding of these factors has the potential to improve understanding reproductive strategies and is, many cases, of direct relevance to management.

@@

>>>>

---
title: |
  Quantifying maternal reproductive output of chondrichthyan fishes
subtitle: |
author:
- name: Alastair V Harry
  affiliation: Fisheries & Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia
  email: alastair.harry@gmail.com
  number: 1
- name: Ivy E Baremore
  affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
  number: 2
- name: Andrew N Piercy
  affiliation: College of Southern Maryland, La Plata MD 20646, United States
  number: 3
toc: false
output:
  word_document: default
  bookdown::pdf_document2:
    pandoc_args:
      - --csl=extras/cjfas.csl
      - --citation-abbreviations=extras/abbreviations.json
      - --template=extras/mytemplate.tex
bibliography: "extras/maternal-output.bib"
keep-latex: true
linestretch: 2 
header-includes:
  - \fontsize{12}{20}
  - \usepackage{lineno}
  - \linenumbers
  - \linenumberfont
  - \renewcommand\linenumberfont{\normalfont\small\sffamily}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{\thepage}
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
editor_options: 
  chunk_output_type: inline
---

@@

**Abstract**

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \> 100 estimation of $P_{Max}$ is generally preferable to using a fixed value.

**Key words** Logistic regression; reproductive biology; maturity ogive; fisheries management

@@

Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass [@hilborn1992; @quinn_quantitative_1999]. Calculating reproductive capacity requires data on the sexually mature proportion of the population contributing to reproduction at a given size or age. For most populations of teleost fishes a simple two-parameter logistic regression function (2PLF) is sufficient to model this process using dichotomous sexual maturity stage data (immature / mature) from a sample of individuals [@jennings_marine_2001; @king_fisheries_2007]. While the mathematical formulae and methods of statistically estimating parameters for a 2PLF vary, the estimated values are similar and can be obtained with the built in generalised linear model (GLM) programs in most modern statistical software.

In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015]. Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017]. This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014]. Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stockrecruitment_2013]. This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

@@

# Materials and methods

@@

A logical progression from estimating $P_{Max}$ is the consideration of alternative functional forms for this parameter such as time- or space-varying $P_{Max}$ or more complex length- or age-dependent forms [@winton2014]. Despite the diverse range of reproductive modes found in chondrichthyan fishes, the current, limited knowledge of maternal investment has typically investigated variables such as litter size and pup condition [@hussey_maternal_2010]. Little is known about how reproductive frequency varies as a function of size or age, in response to environmental effects, or at changing population densities. Better understanding of these factors has the potential to improve understanding reproductive strategies and is, many cases, of direct relevance to management.

@@

# Competing interests

The authors declare there are no competing interests.

# Data availability

Data generated or analyzed during this study are available from the corresponding author upon reasonable request.
